Subset coercions in \Coq

### 2006-07-18 papers/article/main.tex
Minor paper fixes

\input{typing-macros}
\input{typing-decl}
\input{subtyping-decl}

\input{typing-algo}
% \renewcommand{\AppAFull}[9][\Gamma]{
% \TAX{App}
% {$\tgen{#1}{#2}{#3} \suba \Pi #4 : #5. #6$}
% {$\tgen{#1}{#8}{#9}$} % \quad \tgen{#1}{#9, #5}{#7'}$}
% {$\subalgo{#1}{#9}{#5}{s}$}
% {$\tgen{#1}{(#2 #8)}{#6 [ #8 / #4 ]}$}
% {}
% }
\input{subtyping-algo}
\renewcommand{\SubSubAs}[1][\Gamma]{%
\UAX{SubSub}
{$\subalgoimpl{#1}{c}{U}{T}{\Set}$}
{$\subalgoimpl{#1}{c[\eltpit~\ctxdot]}{\mysubset{x}{U}{P}}{T}{\Set}$}
{}
}

@@

The rule \irule{SubConv} integrates $\beta\pi$-conversion in the
judgment. We use a judgemental equality here, which will be refined by
the usual conversion relation later. The \irule{SubSym} and
\irule{SubTrans} rules ensure that our judgement constructs an \emph{equivalence} and has
proper modularity. The next two rules do context closure for dependent
products and sums. It is remarkable that we use contravariance
\cite{journals/toplas/Castagna95} for
domains in the \irule{SubProd} rule, not restricting to invariance as in
\PVS. It is accessory here, as we could have used \emph{co}variance and
still get the same judgements because we have symetry. However, it will
become important when we create coercions (see
\vref{fig:subtyping-impl-rules}).

@@

\begin{paragraph}{}
We use a standard transformation to get an algorithm from our typing
rules, integrating the conversion (or subsumption) rule
(here \irule{Coerce}) to the premises of the other rules.
We recall that $`G \type t : s$ is shorthand for $`G \type t : T$ and
$T  `= s$ with $s `: \PTSsorts$ in a algorithmic presentation.
\TODO{App?, SumDep?, Pi?}. Our algorithmic judgement $`G \typea t : T$
(figure \vref{fig:typing-algo-rules})
reads: $t$ is a well-typed term of type $T$ in environment $`G$.
We need to introduce the notion of support for subset types in order to
define our algorithmic system. Indeed, when typing an application, 
we need to ensure that the object we apply can be seen as an object of a
product type modulo the equivalence. This notion is presented in \PVS{} by
means of a $\mu_0$ operator which was readily adaptable in our setting
\ref{fig:mualgo}. In Gang Chen's work, the notion is formalized by a
new judgement $\Pi_{lub}$ which is essentialy equivalent to the operator.
\begin{figure*}[t]
  \def\infvspace{0.5em}
    \def\type{\typea}
    \def\subt{\subta}
    \def\sub{\suba}
    \def\fCenter{\typea}
    \begin{center}
%     \WfEmpty\DP

@@

@@

\end{paragraph}
\begin{paragraph}{}
  The next thing to do is to construct a decidable judgement for
  equivalence given two types. We denote it by $`G \typea T
  \suba U : s$ (figure \ref{fig:subtyping-algo-rules}).
  In our case, we have to check that conversion is decidable, which is the
  case. We also need to eliminate the transitivity rule which is not
  syntax-directed. That is the purpose of the following theorem:
\begin{theorem}[Elimination of transitivity]
  If $`G \typea T \suba U : s$ and $`G \typea U \suba V : s$, then $`G \typea T
  \suba V : s$.
\end{theorem}
\begin{proof}
  By induction on the sum of the depths of the two derivations.
  \TODO{example?}
\end{proof}
\end{paragraph}

@@

\subsection{Explicit coercions}
The algorithm which derives explicit coercions is based on the
algorithmic equivalence derivations.

>>>>

\input{typing-macros}
\input{typing-decl}
\input{subtyping-decl}
\input{typing-algo}
\input{subtyping-algo}
\renewcommand{\SubSubAs}[1][\Gamma]{%
\UAX{SubSub}
{$\subalgoimpl{#1}{c}{U}{T}{\Set}$}
{$\subalgoimpl{#1}{c[\eltpit~\ctxdot]}{\mysubset{x}{U}{P}}{T}{\Set}$}
{}
}

\input{typing-impl}
\input{subtyping-impl}

@@

The rule \irule{SubConv} integrates $\beta\pi$-conversion in the
judgment. We use a judgemental equality here, which will be refined by
the usual conversion relation later. The \irule{SubSym} and
\irule{SubTrans} rules ensure that our judgement constructs an \emph{equivalence} and has
proper modularity. The next two rules do context closure for dependent
products and sums. It is remarkable that we use contravariance
\cite{journals/toplas/Castagna95} for
domains in the \irule{SubProd} rule, not restricting to invariance as in
\PVS. It is accessory here, as we could have used \emph{co}variance and
still get the same judgements because we have symetry. However, it will
become important when we create coercions (see
\vref{fig:coerce-impl-rules-short}).

@@

\begin{paragraph}{}
We use a standard transformation to get an algorithm from our typing
rules, integrating the conversion (or subsumption) rule
(here \irule{Coerce}) to the premises of the other rules.
We recall that $`G \type t : s$ is shorthand for $`G \type t : T$ and
$T  `= s$ with $s `: \PTSsorts$ in a algorithmic presentation.
\TODO{App?, SumDep?, Pi?}. Our algorithmic judgement $`G \typea t : T$
(figure \vref{fig:typing-algo-rules})
reads: $t$ is a well-typed term of type $T$ in environment $`G$.
We need to introduce the notion of support for subset types in order to
define our algorithmic system. Indeed, when typing an application, 
we need to ensure that the object we apply can be seen as an object of a
product type modulo the equivalence. This notion is presented in \PVS{} by
means of a $\mu_0$ operator which was readily adaptable in our setting
\vref{fig:mualgo-definition}. In Gang Chen's work, the notion is formalized by a
new judgement $\Pi_{lub}$ which is essentialy equivalent to the operator.
\begin{figure*}[t]
  \def\infvspace{0.5em}
    \def\type{\typea}
    \def\subt{\subta}
    \def\sub{\suba}
    \def\fCenter{\typea}
    \begin{center}
%     \WfEmpty\DP

@@

\typemuaFig[$\mualgo$ definition]

@@

\end{paragraph}
\begin{paragraph}{}
  The next thing to do is to construct a decidable judgement for
  equivalence given two types. We denote it by $`G \typea T
  \suba U : s$ (figure \vref{fig:subtyping-algo-rules}).
  In our case, we have to check that conversion is decidable, which is the
  case. We also need to eliminate the transitivity rule which is not
  syntax-directed. That is the purpose of the following theorem:
\begin{theorem}[Elimination of transitivity]
  If $`G \typea T \suba U : s$ and $`G \typea U \suba V : s$, then $`G \typea T
  \suba V : s$.
\end{theorem}
\begin{proof}
  By induction on the sum of the depths of the two derivations.
  \TODO{example?}
\end{proof}
\end{paragraph}

@@

\subsection{Explicit coercions}
The derivation of explicit coercions (figure
\vref{fig:coerce-impl-rules-short}) is based on the
algorithmic equivalence derivations, hence we can derive a function from
the judgement.

\subtqis[Coercion derivation]
